---
title: "Effect on BPE"
author: "Pol + Matei"
format: html
---

#Loadinf all of our data + required packages
```{r}
library(dplyr)
library(here)
BPE_Paris_2016 <- readRDS(here("DATA_Processed" , "BPE_Paris_2016_iris.rds"))
BPE_Paris_2018 <- readRDS(here("DATA_Processed" , "BPE_Paris_2018_iris.rds"))
BPE_Paris_2024 <- readRDS(here("DATA_Processed" , "BPE_Paris_2024_iris.rds"))
Gentri_score <- readRDS(here("DATA_Processed" , "gentrif_indices.rds"))
```

#Making sure that they're the same type 
```{r}
BPE_Paris_2016 <- BPE_Paris_2016 %>%
  mutate(dep = as.character(dep))

BPE_Paris_2018 <- BPE_Paris_2018 %>%
  mutate(dep = as.character(dep))

BPE_Paris_2024 <- BPE_Paris_2024 %>%
  mutate(dep = as.character(DEP))
```

#We're adding the names corresponding to the codes to the 24 base 

```{r}
nomenclature_2024 <- read_html(here("DATA_Raw", "BPE24_liste_hierarchisee_TYPEQU.html"))
   
tables <- html_table(nomenclature_2024, fill = TRUE)

nomenclature_bpe <- tables[[1]] %>%
  clean_names() %>%
  rename(code = code,
         libelle = libelle,
         description = description
         ) %>%
  filter(!is.na(code), code != "")

BPE_Paris_2024 <- BPE_Paris_2024 %>%
  left_join(
    nomenclature_bpe %>% select(code, libelle),
    by = c("TYPEQU" = "code")
  )
```

#Reducing bases into a more compact format and harmonising names at the same time 
```{r}
BPE_Paris_2016_red <- BPE_Paris_2016 %>%
  transmute(
    iris_2020,
    year = 2016,
    code = typequ,
    type = Équipement,
    nb = nb_equip,
    geometry
  )

BPE_Paris_2018_red <- BPE_Paris_2018 %>%
  transmute(
    iris_2020 = CODE_IRIS,
    year = 2018,
    code = code_type_d_equipements,
    type = nom_type_d_equipements,
    nb = nombre_d_equipements,
    geometry
  )

BPE_Paris_2024_red <- BPE_Paris_2024 %>%
  transmute(
    iris_2020,
    year = 2024,
    code = TYPEQU,
    type = libelle,
    nb = 1,   
    geometry
  )

```

After checking BPE_2018_red we notice one code missing it's equipement type code A506 after checking in the following file 
<https://www.google.com/url?sa=t&source=web&rct=j&opi=89978449&url=https://www.insee.fr/fr/metadonnees/source/fichier/BPE1621_nb_equip_UU2020_DictionnaireVariables.pdf&ved=2ahUKEwiDz5v9s_yRAxW0dqQEHVQKAMQQFnoECBoQAQ&usg=AOvVaw3Hhj0EDBeKS9FA4Cg8yyg->

A506 = PRESSING-LAVERIE AUTOMATIQUE ( this code is kept the same in 2024)

```{r}
library(dplyr)

BPE_Paris_2018_red <- BPE_Paris_2018_red %>%
  mutate(
    type = if_else(code == "A506", 
                   "PRESSING-LAVERIE AUTOMATIQUE", 
                   type)
  )
```



#Merging BPE into one base 
```{r}
BPE_final <- bind_rows(
  BPE_Paris_2016_red %>% mutate(year =2016),
  BPE_Paris_2018_red %>% mutate(year = 2018),
  BPE_Paris_2024_red %>% mutate(year = 2024),
)
```



